---
title: "PAg — Professional Agrologist (BC Institute of Agrologists)"
subtitle: "Mapping the UBCO Freshwater Science degree to BCIA's coursework requirements"
date: last-modified
---

::: {.designation-nav}
**Other professional pathways**:
[RPBio (College of Applied Biologists)](rpbio.qmd) ·
[Complementary course bundles (Chemistry, Biology, Hydrology)](complementary.qmd)
:::

# The designation

**PAg** = Professional Agrologist, the senior tier of the
[BC Institute of Agrologists](https://www.bcia.com/). You first qualify
as an **Articling Agrologist (AAg)** straight from your degree, then
complete BCIA's Articling Program plus two years of agrology-related
work experience to bridge to PAg.

## Academic requirements (AAg entry)

A four-year accredited B.Sc. with a science focus, where the coursework
includes:

- **At least 8 entry-level foundational courses** (100/200-level) in
  the subjects on BCIA's Academic Worksheet.
- **At least 20 courses in agriculture, natural sciences, or
  agriculture/resource economics** that relate to agrology under the
  Agrologists Regulation, 2021.
- **At least 8 senior-level courses** (drawn from the 20 above) in
  those same subject areas.

Eligible course categories:

- Agrology
- Foundational Natural Science
- Mathematics or Statistics
- Economics
- Communications / Writing
- Computer Science

# How Freshwater Science maps to PAg

The Freshwater Science core already covers most of BCIA's foundational
requirements. The gap-closers are the senior-level course count and a
small set of category-specific electives.

## What's already covered

- **Foundations (8 required)**: Year 1 science (BIOL 116, CHEM 121 +
  CHEM 123, MATH 100, PHYS 111/112 paired with PHYS 121/122) plus
  **EESC 205** (Hydrology) and **EESC 222** (Geomorphology) cover the
  foundational natural-science count.
- **Communications / Writing**: the program's 6-credit English /
  communications block satisfies BCIA's communications requirement.
- **Mathematics / Statistics**: MATH 100 + STAT 230 / GEOG 271 / BIOL 202.

## BCIA-approved UBCO courses by category {#bcia-approved}

The tables below render directly from the
[June-2025 BCIA UBCO approved course list (PDF)](https://www.bcia.com/sites/default/files/docs/resources/UBCO%20Approved%20Courses_June%2027,%202025.pdf),
parsed into a YAML (`data/designation_mappings/pag.yml`) and committed
to this repository. They update whenever BCIA republishes the list.

```{r}
#| label: pag-tables
#| echo: false
#| warning: false
#| message: false

suppressPackageStartupMessages({
  library(yaml); library(DT)
})
y <- yaml::read_yaml("../data/designation_mappings/pag.yml")

render_category <- function(cat) {
  df <- do.call(rbind, lapply(cat$courses, function(c) {
    data.frame(
      Code  = c$code,
      Title = c$title,
      `Cross-listed` = if (length(c$cross_codes))
        paste(unlist(c$cross_codes), collapse = ", ") else "",
      check.names = FALSE,
      stringsAsFactors = FALSE
    )
  }))
  DT::datatable(
    df, rownames = FALSE, escape = FALSE,
    options = list(pageLength = 15, dom = "tip",
                   columnDefs = list(list(className = "dt-left",
                                          targets = "_all")))
  )
}
```

### Foundational courses (8 required, 100/200-level)

`r length(yaml::read_yaml("../data/designation_mappings/pag.yml")$categories$foundational$courses)` UBCO courses are recognised in this category.

```{r}
#| echo: false
render_category(y$categories$foundational)
```

### Agrology courses — 100/200 level

`r length(yaml::read_yaml("../data/designation_mappings/pag.yml")$categories$agrology_lower$courses)` UBCO courses in the lower-level agrology category. These can count toward the 20-course agrology requirement (≥8 must be senior, listed below).

```{r}
#| echo: false
render_category(y$categories$agrology_lower)
```

### Agrology courses — 300/400+ level

`r length(yaml::read_yaml("../data/designation_mappings/pag.yml")$categories$agrology_senior$courses)` UBCO senior-level agrology courses. BCIA requires **≥8 senior-level courses** within the 20-course agrology total.

```{r}
#| echo: false
render_category(y$categories$agrology_senior)
```

## Sample 4-year plan for the PAg track

Illustrative; verify with the program advisor and current BCIA rules.

::: {.callout-note collapse="true"}
## Year 1
- **T1**: BIOL 116; CHEM 121; MATH 100; ENGL 109/112/114 (writing requirement); elective.
- **T2**: BIOL 125; CHEM 123/121; PHYS 111 or 112; elective.
:::

::: {.callout-note collapse="true"}
## Year 2
- **T1**: EESC 205 (Hydrology); STAT 230 / GEOG 271 / BIOL 202; ECON 371
  or 372 (example); elective.
- **T2**: EESC 222 (Geomorphology); CHEM 211; CHEM 201 or 210
  (example); elective.
:::

::: {.callout-note collapse="true"}
## Year 3
- **T1**: EESC 301 (Limnology); EESC 402 or other policy/management
  upper-level; GISC/GEOG 380/381 (GIS/Remote Sensing).
- **T2**: Choose among EESC 333, 413, 422, 429, 431, 435; consider CHEM
  301/302 as relevant.
:::

::: {.callout-note collapse="true"}
## Year 4
- Round out to **≥20 agrology-related courses with ≥8 at 3rd/4th-year**
  across EESC, BIOL, CHEM, GEOG, GISC, ECON. Add a field or industry
  project (Directed Studies, Honours Thesis, or co-op term).
:::

## Co-op and experience

Co-op terms after Year 2 or Year 3 target utilities, local and regional
governments, environmental consulting, and agriculture / watershed
NGOs. These count toward the post-degree articling experience required
to convert AAg → PAg.

## Seat / prerequisite cautions

- GIS courses (GISC 380/381) often have priority rules for majors.
  Request permission early.
- ECON seats fill quickly; plan registration day.
- Some upper-level EESC courses alternate years (313/413). Coordinate
  with the program advisor.

# Your gap-finder (interactive)

::: {.callout-tip}
## When the planner is wired up
The interactive planner will read your completed-courses list and the
authoritative BCIA UBCO course mapping and show exactly which BCIA
categories are covered and which still need fills.
:::

# References

- [New Registrant Requirements (BCIA)](https://www.bcia.com/how-apply/application-requirements/new-registrant-requirements)
- [BCIA UBCO Approved Course List (June 2025 PDF)](https://www.bcia.com/sites/default/files/docs/resources/UBCO%20Approved%20Courses_June%2027,%202025.pdf)
- [Agrologists Regulation, 2021](https://www.bclaws.gov.bc.ca/civix/document/id/complete/statreg/47_2021)
- Source notes in this repo: `docs/reference/advising/UBCO_Freshwater_Professional_Paths_Advising_Pack.pdf`.
