---
title: "RPBio — Registered Professional Biologist (College of Applied Biologists)"
subtitle: "Mapping the UBCO Freshwater Science degree to CAB's coursework requirements"
date: last-modified
---

# The designation

**RPBio** = Registered Professional Biologist, granted by the
[College of Applied Biologists](https://cab-bc.org/) — the regulatory
body for applied biology professionals in BC. Entry-level is
**Biologist in Training (BIT)** with the academic credential complete;
RPBio follows after validated experience, competencies, and submitted
work products.

## Academic requirements

A bachelor's degree (or master's) with:

- **25 science courses, 13 in biology.**
- Required core courses across:
  - **Communications** (typically first-year writing)
  - **Chemistry**
  - **Numeracy** (math)
  - **Statistics** at 2nd-year or higher
  - **Applied biology / management**
  - **Ecology**
  - **Three of five** from: **genetics, cell biology, physiology,
    systematics/taxonomy, evolution**

::: {.callout-warning}
## Known UBCO gap: systematics / taxonomy
UBCO does not currently offer a dedicated undergraduate course in
systematics or taxonomy. Students targeting RPBio should either
(a) plan to satisfy this requirement via a course taken at another
institution and transferred in, or (b) cover the three-of-five via
**evolution + cell biology + genetics** (or + physiology) instead.
**Confirm with the program advisor and CAB before registering.**
This gap was identified during the May 2026 QAQC pass against the
live UBC Okanagan Academic Calendar.
:::

# How Freshwater Science maps to RPBio

The Freshwater Science core already covers most of the chemistry,
numeracy, statistics, applied-biology, and ecology requirements. The
gap to close is the biology course count (need 13) and the
"three-of-five subject areas" requirement.

## What's already covered

- **Communications**: first-year writing block (ENGL or
  program-approved alternatives).
- **Chemistry**: CHEM 121/123, 201/210/211, 301 — comfortable coverage.
- **Numeracy**: MATH 100/101.
- **Statistics (≥2nd-year)**: STAT 230 or GEOG 271 or BIOL 202.
- **Ecology / applied biology**: BIOL 201 (Ecology & Evolution),
  BIOL 308 (Population Biology), FWSC/BIOL 375 (Flora & Fauna of
  Inland Waters).

## Required core areas + the three-of-five subject areas

The tables below render directly from
`data/designation_mappings/rpbio.yml` — the curated mapping of CAB's
requirements onto UBCO courses, based on the
[CAB Credentialing Standard (June 2025)](https://cab-bc.org/wp-content/uploads/Credentialing_Standard_June-2025.pdf)
plus the program-lead's advising pack (CAB does not publish a
UBCO-specific approved-course PDF).

```{r}
#| label: rpbio-tables
#| echo: false
#| warning: false
#| message: false

suppressPackageStartupMessages({
  library(yaml); library(DT)
})
y <- yaml::read_yaml("../data/designation_mappings/rpbio.yml")

render_area <- function(area) {
  df <- do.call(rbind, lapply(area$courses_eligible, function(c) {
    data.frame(
      Code  = c$code,
      Title = c$title,
      check.names = FALSE,
      stringsAsFactors = FALSE
    )
  }))
  DT::datatable(
    df, rownames = FALSE, escape = FALSE,
    options = list(pageLength = 8, dom = "t",
                   columnDefs = list(list(className = "dt-left",
                                          targets = "_all")))
  )
}
```

### Required core areas

These six areas all need at least one qualifying UBCO course on the
transcript.

#### Communications
```{r}
#| echo: false
render_area(y$categories$communications)
```

#### Chemistry
```{r}
#| echo: false
render_area(y$categories$chemistry)
```

#### Numeracy
```{r}
#| echo: false
render_area(y$categories$numeracy)
```

#### Statistics (≥ 2nd-year)
```{r}
#| echo: false
render_area(y$categories$statistics_second_year_or_higher)
```

#### Applied biology / management
```{r}
#| echo: false
render_area(y$categories$applied_biology_management)
```

#### Ecology
```{r}
#| echo: false
render_area(y$categories$ecology)
```

### Three-of-five subject-area requirement

CAB requires at least **three of these five** subject areas to be
covered. Evolution (BIOL 201) is already in the Freshwater Science
core, so the practical target is "two more of genetics / cell biology /
physiology / systematics".

#### Genetics
```{r}
#| echo: false
render_area(y$categories$genetics)
```

#### Cell biology
```{r}
#| echo: false
render_area(y$categories$cell_biology)
```

#### Physiology
```{r}
#| echo: false
render_area(y$categories$physiology)
```

#### Systematics / taxonomy
```{r}
#| echo: false
render_area(y$categories$systematics_taxonomy)
```

#### Evolution
```{r}
#| echo: false
render_area(y$categories$evolution)
```

### Biology-count check

CAB requires **13 biology courses**. The planner uses this list of
UBCO codes when computing your biology-count coverage:

```{r}
#| echo: false
codes <- y$biology_count_eligible_codes
cat("**", length(codes), " UBCO course codes** counted toward the 13-biology threshold:\n\n",
    paste0("`", codes, "`", collapse = "  ·  "), sep = "")
```

## Sample 4-year plan for the RPBio track

Illustrative; verify with the program advisor and current CAB rules.

::: {.callout-note collapse="true"}
## Year 1
- **T1**: BIOL 116; CHEM 121; MATH 100; ENGL 109/112/114 (writing requirement); elective.
- **T2**: BIOL 125; the other of CHEM 121/123; PHYS 111 or 112;
  elective.
:::

::: {.callout-note collapse="true"}
## Year 2
- **T1**: BIOL 201; STAT 230 / GEOG 271 / BIOL 202 (≥2nd-year stats);
  EESC 205 (Hydrology); elective.
- **T2**: EESC 222 (Geomorphology); CHEM 211; **BIOL 200 (Cell)**;
  elective.
:::

::: {.callout-note collapse="true"}
## Year 3
- **T1**: EESC 301 (Limnology); FWSC/BIOL 375 (Flora & Fauna of Inland
  Waters); CHEM 301; elective.
- **T2**: **BIOL 265 (Genetics)**; **BIOL 356 or 358 (Physiology)** or
  BIO upper-level toward the 13-biology count.
:::

::: {.callout-note collapse="true"}
## Year 4
- **T1**: EESC 402 (Freshwater Resources); pick 1-2 of EESC 309 / 323 /
  423 (geochem requirement); BIO upper-level (reaching the 13-BIO total).
- **T2**: EESC 342 (Hydrogeology) or EESC 413 (Analytical Methods); BIO
  upper-level; Directed Studies / Field / Special Topics.
:::

## Co-op and experience

Co-op terms after Year 2 (summer start) for 8-12 months, or after
Year 3 summer. Target biomonitoring, fisheries, environmental
consulting, and government. Experience accumulated during BIT counts
toward the RPBio bridge.

## Seat / prerequisite cautions

- Some 3xx BIOL seats prioritise Biology majors. Request permission
  early.
- BIOL 200, 265, 356, 358, 371 are the gap-filler electives — plan
  registration so they aren't squeezed by required EESC courses.

## Professional work products

To convert BIT → RPBio you must submit at least one **scientific or
risk-assessment report** that meets College criteria, plus at least one
**public-facing communication piece** (annotated presentation, briefing
note, interpretation document, etc.). Directed Studies, Honours Thesis,
and co-op term reports are the natural vehicles for these.

# Your gap-finder (interactive)

::: {.callout-tip}
## When the planner is wired up
The interactive planner cross-checks your completed-courses list
against the CAB biology categories and flags which of the five
required-area courses (genetics, cell, physiology, systematics,
evolution) you have versus still need.
:::

# References

- [RPBio Requirements (CAB)](https://cab-bc.org/applicant-self-assessment-tool-general-information/applicant-self-assessment-tool/rpbio-form/)
- [Credentialing Standard (June 2025 PDF)](https://cab-bc.org/wp-content/uploads/Credentialing_Standard_June-2025.pdf)
- [College of Applied Biology Act](https://www.bclaws.gov.bc.ca/civix/document/id/consol14/consol14/00_02068_01)
- Source notes in this repo: `docs/reference/advising/UBCO_Freshwater_Professional_Paths_Advising_Pack.pdf`.
